Illuminating a Dark Lens : A Type Ia Supernova Magnified by \\ the Frontier Fields Galaxy Cluster Abell 2744

### 2016-04-07 APJ.accepted/response.txt

2015.08.10 
Author's Response to Referee Report

Dear Editors, 
 The referee provided a very positive and helpful report with only minor changes required.  Below, we provide the referee's comments (REF) and our responses (SR).   The re-submitted manuscript highlights all additions or changes with blue text.  Thank you for your attention to this paper, and please let us know if anything else is needed.

Sincerely,
Steve Rodney and co-authors

REF:   This paper reports the discovery of a type Ia supernova behind one of the HFF clusters, and uses its measured magnification to test cluster mass models. The discovery is definitely important, and this indeed provides quite a rare opportunity to do a blind test of many mass models constructed by different groups. The analysis is conducted very carefully to get convincing results, and the results are also presented very well. This paper clearly deserves publication in ApJ. I only have a few minor comments which I think are straightforward to address.

-----------------

REF:   1. HFF14Tom is located outside the strong lensing region, so naive expectation is that adding weak lensing constraints more or less reduces the potential bias in predicting its magnification. I'm sure the authors are aware of this, as they use different symbols for methods combining strong and weak lensing in e.g., Figure 6, but there is virtually no discussion about the role of weak lensing in the text. The SN region can at least be probed well by HST weak lensing. Even ground based weak lensing can get reasonable constraints on enclosed mass within 1 arcmin radius or so if the SN is high. On the other hand, looking at Figure 6 it is not clear if methods using weak lensing are better than strong lensing only methods. In particular CATS v1.1 is worse than CATS v1, which looks a bit strange given that their difference come from weak lensing constraints. Anyways, I think it's worth adding a paragraph or two to discuss effects of weak lensing constraints. I'm interested in the authors' opinion about this.   Also please describe (e.g., in Table 4) which weak lensing data each method used.

SR: A discussion of strong vs strong+weak lensing has been added to section 7 when first introducing the models.  Additionally, a new subsection (7.2.1) discusses how weak lensing could possibly avoid a particular extrapolation bias, but we conclude that there is no evidence for such a bias in this case.   Also, Table 4 has been updated with new table notes indicating the weak lensing data used.

-----------------

REF:  2. Sec. 7.2.5: Are there any good references for the discussion in the first paragraph? I see the point that the source plane minimization prefers high magnification solutions, but this is the case only when distances in the source plane (I mean the distances between average source position and source positions computed from individual image positions) are naively minimized. In fact one can convert these distances to corresponding image plane distances using magnifications (or more rigorously the Jacobian matrix), and once these corrections are applied this bias can easily be avoided. I know several methods (such as LENSTOOL, GLAFIC, ...) already implement this, although some details look different between different codes. It may be worth studying further which models are possibly affected by this source plane minimization bias.

SR:  We have updated this section (now 7.2.6) with some additional references, and clarifying text.   The Kochanek 1991 and Kochanek 2006 references provide a more general discussion of the issues of source plane vs image plane minimization.  Jee et al 2007 and Ponente & Diego 2011 include sections dealing more specifically with the bias toward high magnification.

-----------------

REF:  3. Figure 8 is nice, but it's a bit unclear what the y-axis means. It takes negative values, so it should be something like (\mu_model-\mu_SN)/\sigma, but \sigma here includes only the measurement error on the magnification from HFF14Tom, or does it also include model errors listed in Table 5? If so those errors are quadratically added?

SR:  Yes.  The referee is precisely correct on all counts.  This has been added in the figure caption.

-----------------

REF:  4. Sec. 5: ``we do not include any prior on...' Do you mean by flat prior?

SR:  Yes.  The text has been updated.

-----------------

REF:  5. Sec. 5: ``... the host galaxy photo-z and spectroscopic redshift (Section 2 and 4.2)'. I believe it is Section 3 not Section 2. In the caption of Figure 3 ``consistent with both the photo-z of the probable...' it should be spectroscopic redshift rather than photo-z.

SR: These corrections have been made.    We have also made a few other small corrections in the text that were not noted by the referee.  These are also highlighted in blue text in the revised manuscript.

### 2016-04-07 APJ.accepted/authors.tex
updated files for accepted version

@@

@@

\newcommand{\CalTech}{California Institute of Technology, 1200 East California Boulevard, Pasadena, CA 91125}

@@

\author{Steven~A.~Rodney\altaffilmark{\affilref{JHU},\affilref{HubbleFellow}}}
\affilreftxt{JHU}{\JHU}
\affilreftxt{HubbleFellow}{\HubbleFellow}

@@

\author{Daniel~Scolnic\altaffilmark{\affilref{Chicago}}}
\affilreftxt{Chicago}{\Chicago}

@@

\author{Adam G.~Riess\altaffilmark{\affilref{JHU},\affilref{STScI}}}
\affilreftxt{JHU}{\JHU}
\affilreftxt{STScI}{\STScI}

>>>>

\newcommand{\USC}{Department of Physics and Astronomy, University of South Carolina, 712 Main St., Columbia, SC 29208, USA}

@@

\newcommand{\Kavli}{Kavli Institute for Cosmological Physics, University of Chicago, 5640 South Ellis Avenue Chicago, IL 60637}

@@

\newcommand{\CalTech}{Cahill Center for Astronomy and Astrophysics, California Institute of Technology, MC 249-17, Pasadena, CA 91125, USA}

@@

\author{Steven~A.~Rodney\altaffilmark{\affilref{USC},\affilref{JHU},\affilref{HubbleFellow}}}
\affilreftxt{USC}{\USC}
\affilreftxt{JHU}{\JHU}
\affilreftxt{HubbleFellow}{\HubbleFellow}
\email{srodney@sc.edu}

@@

\author{Daniel~Scolnic\altaffilmark{\affilref{Kavli}}}
\affilreftxt{Kavli}{\Kavli}

@@

\author{Adam Riess\altaffilmark{\affilref{JHU},\affilref{STScI}}}
\affilreftxt{JHU}{\JHU}
\affilreftxt{STScI}{\STScI}

### 2016-04-07 APJ.accepted/lensmodels.tex
updated files for accepted version

CATS(v2)       & 50 & 151 & 4 & 1  &  \citealt{Jauzac:2014c} & Updated version of the CATSv1 model, adds 33 new multiply-imaged galaxies, for a total of 159 individual lensed images.\\

@@

CATS(v2.1)     & 55 & 154 & 8 & 1  &  \citealt{Jauzac:2014c} & Updated version of the CATSv2 model, adopting the spec-z constraints used for Bradac(v2).\\
CATS(v2.2)     & 25 &  72 & 8 & 1  &  \citealt{Jauzac:2014c} & Updated version of the CATSv2 model, adopting the spec-z constraints and multiple-image definitions used for Bradac(v2).

>>>>

CATS(v2)       & 50 & 151 & 4 & 1  &  \citealt{Jauzac:2015b} & Updated version of the CATSv1 model, adds 33 new multiply-imaged galaxies, for a total of 159 individual lensed images.\\

@@

CATS(v2.1)     & 55 & 154 & 8 & 1  &  \citealt{Jauzac:2015b} & Updated version of the CATSv2 model, adopting the spec-z constraints used for Bradac(v2).\\
CATS(v2.2)     & 25 &  72 & 8 & 1  &  \citealt{Jauzac:2015b} & Updated version of the CATSv2 model, adopting the spec-z constraints and multiple-image definitions used for Bradac(v2).

### 2016-04-07 APJ.accepted/snTomas.tex
updated files for accepted version

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%  PREAMBLE: sets up compiler modes, loads packages, defines macros, etc
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\input{preamble}

@@

In addition to time delays from multiply-imaged SNe, we can also
put cluster mass models to the test with the much more common category
of Type Ia SNe that are magnified but not multiply-imaged.  
\citet[][hereafter P14]{Patel:2014} and \citet{Nordin:2014} presented
independent analyses of three lensed SNe, of which at least 2 are
securely classified as Type Ia SNe -- all found in the Cluster Lensing
and Supernova survey with Hubble (CLASH, PI:Postman, HST Program ID
12068, \citealt{Postman:2012}).  Both groups demonstrated that these
standard candles can be used to provide accurate and precise
measurements of the true absolute magnification along a random sight
line through the cluster.  Although in these cases the SNe were used
to {\it test} the cluster mass models, one could in principle
incorporate the measured magnifications of Type Ia SNe into the
cluster as additional model constraints.  In that role, Type Ia SNe
have the particular value that they can be found anywhere in the
cluster field.  Thus, they can deliver model constraints in regions of
``middle distance'' from the cluster core, where both strong- and
weak-lensing constraints are unavailable. Moreover, given
enough time, multiple background SNe Ia could be measured behind the
same cluster, each providing a new magnification constraint.

@@

\begin{figure*}
\begin{center}
%\includegraphics[width=\textwidth]{discovery_image}
\includegraphics[width=\textwidth]{discovery_image_lowres}
\caption{  \label{fig:DiscoveryImage} 
SN \tomas\ in the Abell 2744 field.  The left panel shows a
UV/Optical/IR color composite image constructed from all available HST
imaging of the Abell 2744 cluster field.  The inset panels on the right
show F814W imaging of the immediate vicinity of SN \tomas,
approximately 40\arcsec\ from the center of the cluster. The top panel
shows the template image, combining all data prior to the SN
appearance.  Labeled ellipses mark the nearest galaxies and their
spectroscopic redshift constraints, with the most likely host
galaxy marked in blue, and a background galaxy in yellow. The bottom
panel is constructed from all HFF F814W imaging taken while the SN was
detectable, and marks the SN location with an arrow.  (Left panel
image credit: NASA, ESA, and J. Lotz, M. Mountain, A. Koekemoer, and
the HFF Team (STScI)) }
\end{center}
\end{figure*}

@@

From the preceding analysis, we conclude that \tomas\ is a Type Ia SN
at $z\approx1.35$.  At this redshift, the absorption at $\sim$8700\AA\
corresponds to the blended \ion{Ca}{2} H\&K and \ion{Si}{2} $\lambda$3858 features.  Generally referred to as the Ca H\&K feature, this absorption is commonly seen in Type Ia SN spectra
near maximum light, although it is also prominent in the spectra of
Type Ib and Ic core collapse SNe (\CCSNe).  The red color of
the \tomas\ SED is qualitatively consistent with a redshift of $z>1$
-- although this information was not used by {\tt SNID} for the
template matching.  As we will see in
Section~\ref{sec:PhotometricClassification}, this spectral
classification of SN \tomas\ is reinforced by the photometric
information, which also supports classification as a Type Ia SN at
$z\approx1.35$.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%  PREAMBLE: sets up compiler modes, loads packages, defines macros, etc
%  Steve Rodney, 2012
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%
% COMPILER MODES
%%%%%%%%%%%%%%%%%%%%

%% Set default compiler modes 
%% (these will be overridden by the -options.tex  file below, if it exists)

% Changetext mode : highlight modified text in bold blue font
\newif{\ifchangetext}
%\changetexttrue
\changetextfalse

\documentclass[iop]{emulateapj}
\citestyle{aa}

%%%%%%%%%%%%%%%%%%%%
% PACKAGES INCLUDED
%%%%%%%%%%%%%%%%%%%%
\usepackage{ifthen}
\usepackage{natbib}   % reference citations and bibliography
\usepackage{amsmath}  % extended math symbols
\usepackage{hyperref}      % url text formatting
\usepackage[usenames,dvipsnames]{xcolor}  % colored text
%\usepackage{pdflscape}

%\usepackage{amssymb}  % extended symbols lib
%\usepackage{verbatim} % verbatim text formatting
%\usepackage{enumerate}% enumerated lists
%\usepackage{url}      % url text formatting
%\usepackage{multirow}  % muti-row table cells
%\usepackage{amsmath}  % extended equations lib (split)
%\usepackage{mathrsfs} % extended math fonts (mathscr)
%\usepackage{paralist} % inline enumeration (for Table ref lists)
%\usepackage{authblk}
%\usepackage{appendix}
%\usepackage{array}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% changetext  mode settings
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\ifchangetext
  % Changed text is highlighted in bold, blue font 
  \newcommand{\change}[1]{{ \textcolor{blue}{#1}}}
  \newcommand{\changenote}[1]{{ \textcolor{blue}{#1}}}
  \newcommand{\changeB}[1]{{ \textcolor{blue}{#1}}}
  \newcommand{\changeG}[1]{{ \textcolor{ForestGreen}{#1}}}
  \newcommand{\changeR}[1]{{ \textcolor{red}{#1}}}
  \newcommand{\changeP}[1]{{ \textcolor{DarkOrchid}{#1}}}
\else
  % Changed text is indistinguishable
  \newcommand{\change}[1]{#1}
  \newcommand{\changeB}[1]{#1}
  \newcommand{\changeG}[1]{#1}
  \newcommand{\changeR}[1]{#1}
  \newcommand{\changeP}[1]{#1}
  \newcommand{\changenote}[1]{}
\fi

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% PDF mode settings : Auto-select eps or pdf figures 
% based  on the compiler used (i.e. latex vs pdflatex)
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\DeclareGraphicsExtensions{.pdf,.png,.jpg}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% AUTHOR-DEFINED MACROS
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Paper aliases 
\defcitealias{Patel:2014}{P14}
\def\P14{\citetalias{Patel:2014}}

\def\tomas{HFF14Tom}

% Cosmology:
\def\Om{\ensuremath{\Omega_{\rm m}}}
\def\Ot{\ensuremath{\Omega_{\rm tot}}}
\def\Ob{\ensuremath{\Omega_{\rm b}}}
\def\OL{\ensuremath{\Omega_{\Lambda}}}
\def\Ok{\ensuremath{\Omega_{\rm k}}}
\def\om{\ensuremath{\omega_{\rm m}}}
\def\ob{\ensuremath{\omega_{\rm b}}}
\def\wo{\ensuremath{w_0}}
\def\wa{\ensuremath{w_{\rm a}}}
\def\lcdm{$\Lambda$CDM}
\def\LCDM{$\Lambda$CDM}
\def\wcdm{$w$CDM}
\def\Ho{\ensuremath{H_0}}
\def\DA{\ensuremath{D_A}}
\def\DL{\ensuremath{D_L}}

% Astronomy:
\def\arcsec{\ensuremath{^{\prime\prime}}} 
\def\kms{\ensuremath{{\rm km s}^{-1}}}
\def\hgpcq{\mbox{$h^{-3}$Gpc$^3$}}
\def\hmpcq{\mbox{$h^{-3}$Mpc$^3$}}
\def\perhmpcq{\mbox{$h^{3}$Mpc$^{-3}$}}
\def\hmpc{\mbox{$h^{-1}$Mpc}}
\def\hmpci{\mbox{$h$\,Mpc$^{-1}$}}
\def\mpc{\mbox{Mpc}}
\def\mpci{\mbox{Mpc$^{-1}$}}
\def\mpcq{\mbox{Mpc$^{-3}$}}
\def\Msun{\mbox{M$_{\odot}$}}
\def\Av{\mbox{$A_V$}}
\def\Rv{\mbox{$R_V$}}

% Supernovae : 
\newcommand{\CCSN}{core collapse SN}
\newcommand{\CCSNe}{core collapse SNe}
\newcommand{\SNIa}{Type Ia SN}
\newcommand{\SNeIa}{Type Ia SNe}
\newcommand{\SNRz}{SNR($z$)}
\def\Mch{\mbox{M$_{\rm Ch}$}}
\def\Ni{\ensuremath{^{56}\mbox{Ni}}}
\newcommand{\dmfifteen}{\ensuremath{\Delta\mbox{m}_{15}}}
\newcommand{\deltamfifteen}{\ensuremath{\Delta\mbox{m}_{15}}}

% Missions:
\def\HST{{\it HST}}
\def\Hubble{{\it Hubble}}
\def\Hubbles{{\it Hubble's}}
\def\Spitzer{{\it Spitzer}}
\def\Chandra{{\it Chandra}}
\def\Herschel{{\it Herschel}}
\def\XMM{{\it XMM}}

@@

In addition to time delays from multiply-imaged SNe, we can also
put cluster mass models to the test with the much more common category
of Type Ia SNe that are magnified but not multiply-imaged.  
\citet[][hereafter P14]{Patel:2014} and \citet{Nordin:2014} presented
independent analyses of three lensed SNe, of which at least 2 are
securely classified as Type Ia SNe -- all found in the Cluster Lensing
and Supernova survey with Hubble (CLASH, PI:Postman, HST Program ID
12068, \citealt{Postman:2012}).  Both groups demonstrated that these
standard candles can be used to provide accurate and precise
measurements of the true absolute magnification along a random sight
line through the cluster.  Although in these cases the SNe were used
to {\it test} the cluster mass models, one could in principle
incorporate the measured magnifications of Type Ia SNe into the
cluster as additional model constraints.  In that role, Type Ia SNe
have the particular value that they can be found anywhere in the
cluster field.  Thus, they can deliver model constraints in regions of
``middle distance'' from the cluster core, where both strong- and
weak-lensing constraints are unavailable. Moreover, given
enough time, multiple background \SNeIa\ could be measured behind the
same cluster, each providing a new magnification constraint.

@@

\begin{figure*}
\begin{center}
%\includegraphics[width=\textwidth]{discovery_image_lowres}
\includegraphics[width=\textwidth]{discovery_image_medres}
%\includegraphics[width=\textwidth]{discovery_image_highres}
\caption{  \label{fig:DiscoveryImage} 
SN \tomas\ in the Abell 2744 field.  The left panel shows a
UV/Optical/IR color composite image constructed from all available HST
imaging of the Abell 2744 cluster field.  The inset panels on the right
show F814W imaging of the immediate vicinity of SN \tomas,
approximately 40\arcsec\ from the center of the cluster. The top panel
shows the template image, combining all data prior to the SN
appearance.  Labeled ellipses mark the nearest galaxies and their
spectroscopic redshift constraints, with the most likely host
galaxy marked in blue, and a background galaxy in yellow. The bottom
panel is constructed from all HFF F814W imaging taken while the SN was
detectable, and marks the SN location with an arrow.  (Left panel
image credit: NASA, ESA, and J. Lotz, M. Mountain, A. Koekemoer, and
the HFF Team (STScI)) }
\end{center}
\end{figure*}

@@

From the preceding analysis, we conclude that \tomas\ is a Type Ia SN
at $z\approx1.35$.  At this redshift, the absorption at $\sim$8700\AA\
corresponds to the blended \ion{Ca}{2} H\&K and \ion{Si}{2} $\lambda$3858 features.  Generally referred to as the Ca H\&K feature, this absorption is commonly seen in Type Ia SN spectra
near maximum light, although it is also prominent in the spectra of
Type Ib and Ic \CCSNe.  The red color of
the \tomas\ SED is qualitatively consistent with a redshift of $z>1$
-- although this information was not used by {\tt SNID} for the
template matching.  As we will see in
Section~\ref{sec:PhotometricClassification}, this spectral
classification of SN \tomas\ is reinforced by the photometric
information, which also supports classification as a Type Ia SN at
$z\approx1.35$.
